The role of advection versus diffusion in intraventricular transport

### 2025-03-06 paper/main.tex
Minor updates in experimental methods

The larval and adult zebrafish were raised under standard husbandry
conditions at 28.5°C in a Techniplast Zebtech Multilinking system.
The fish tanks were kept at constant pH 7.0 and 685 \textmu S with a
14/10 hr light/dark cycle.
From fertilization to 3 dpf, larvae were maintained in egg
water (1.2 g marine salt and 0.1\% methylene blue in 20 L reverse osmosis (RO) water)
and subsequently transferred to artificial fish water (AFW)
(1.2 g marine salt in 20 L RO water). The zebrafish lines used in the experiments were
\emph{ccdc103(dnaaf19)\textsuperscript{tn222a}}
(\emph{schmalhans (smh)})~\cite{Jau-NianChen1997Left-rightZebrafish}
and \emph{Tg(gfap:Gal4FF)\textsuperscript{nw7Tg}}~\cite{DiazVerdugo2019Glia-neuronSeizures},
\emph{Tg(5xuas:Signal-Dendra2)\textsuperscript{nw21Tg}}.
Animals were in the pigmentless \emph{nacre\textsuperscript{b692}}
(\emph{mitfa\textsuperscript{-/-}})~\cite{JamesA.Lister1999NacreFate} background.

Wholemount in vivo live imaging and photoconversion
experiments were performed with zebrafish larvae at 2 days post-fertilization (dpf) stage obtained from inbreeding of heterozygous
\emph{smh\textsuperscript{+/-};Tg(gfap:Gal4FF);Tg(5xuas:Signal-Dendra2})
adult animals, as described in D'Gama~\emph{et al.}~\cite{DGama2024Cilia-mediatedBrain}.
Controls consisted of either wild-type (\emph{smh}\textsuperscript{+/+})
or \emph{smh}\textsuperscript{+/-} heterozygous zebrafish from the same breeding.
The mutants were identified based on their curved body~\cite{Jau-NianChen1997Left-rightZebrafish}.

For genotyping, genomic DNA (gDNA) was isolated from clipped fins of anesthetized adult
fish using 100 \textmu L of PCR lysis buffer (containing 1M tris pH 7--9, 0.5 M EDTA,
tritonX-100, and Proteinase K 0.1 mg/ml) overnight at 50°C.
To stop the lysis reaction, the samples were heated to 95°C for 10 minutes
and then centrifuged at 13000 rpm for 2 minutes.
The supernatant containing gDNA was utilized for KASP assays-based analysis.
The gDNAs were diluted (1:2) with water, and 3 \textmu L of the diluted gDNA was
used to perform the KASP assay following the guidelines of the manufacturer (LGC Biosearch Technologies™).
For each sample well, the master mix contained: 5 \textmu L of master mix; 0.14
\textmu L of assay mix; 1.86 \textmu L of RO water.

@@

\subsection{Wholemount zebrafish in vivo live time-lapse imaging and photoconversion of Dendra2}
Two dpf larval zebrafish were anesthetized in 0.013\% MS222
(Ethyl 3-aminobenzoate methanesulfonate, Sigma) in AFW and
mounted laterally in 1.5\% low melting agarose in a Flurodish (VWR, FD35PDL-100),
to obtain a lateral view of the brain ventricles. After the
fish were positioned in the agarose, the dish stood for 5 min to allow the agarose to solidify.
After solidification, AFW containing 0.013\% MS222 was added,
and then the fish were transferred to the confocal microscope (LSM880 Examiner, Zeiss)
and imaged using a 20x water-immersion objective (Zeiss, NA 1.0, Plan-Apochromat)
at room temperature.

@@

\subsection{Photoconversion data analysis}
All data analysis was performed in MATLAB and the codes are available in
the repository of this article~\cite{zenodo-link-to-all-code}\fixme{Add repo link.}.

Following import of the image data, the photoconverted channel
was aligned to correct for drift in $x,y$ directions using previously
developed codes~\cite{Reiten2017Motile-Cilia-MediatedComputations, Ringers2023NovelEpithelia}.
Only stable recordings were further analyzed. 
To increase the signal to noise ratio, the images were downsampled by
a block average using a resampling factor of 6 (using the MATLAB function blockproc).
The change of fluorescence $($$(F-F_0)/F_0$, hereafter referred to as
$\Delta F$$)$ was calculated for each pixel at each time point of the data.
Here, the value $F_0$ is the average fluorescence intensity for the
baseline (the first 10 time-lapse images) acquired before photoconversion.

@@

To obtain $\Delta F$ values for a region of interest (ROI), ROIs were
first drawn on the aligned and block-averaged time series.
Six ROIs were drawn manually: ROI 1 around the location of photoconversion,
and ROIs 2--6 in different regions of the ventricular system.
The pixel values of the fluorescence intensity within one ROI were
averaged for each time point, and the relative change $\Delta F$ calculated with this averaged value.
To normalize the data with respect to photoconversion efficiency, the
$\Delta F$ curves for all ROIs were divided by the $\Delta F$ values
obtained for ROI 1 at the end of the experiment,
so that $\Delta F$ values for the photoconverted site ROI 1 ranged from
zero to one. Finally, we averaged the $\Delta F$ data from controls and from mutants.
We report these mean values together with a one-standard deviation error band.

\subsection{Intraventricular injection of microbeads and particle tracking}
To recover the flow fields shown in~\Crefpanels{fig:fig1}{b, c}, we carried out
particle tracking in a larval zebrafish injected with fluorescent beads with a procedure described in detail in previous work~\cite{Olstad2019CiliaryDevelopment}.
Anesthetized 2 dpf zebrafish were injected with 1 nl of a mixture containing 0.1\% w/v fluorescent beads
(SPHERO Fluorescent Yellow Particles 1\% w/v, F = 0.16 mm) diluted
in 7.5 mg/ml 70 kDa rhodamine B isothiocyanate-dextran (RITCdextran; Sigma-Aldrich, R9379)
dissolved in artificial CSF. Artificial CSF composition was as follows:
124 mM NaCl, 22 mM D-(+)-Glucose, 2.0 mM KCl, 1.6 mM MgSO\textsubscript{4} $\cdot$ 7 H\textsubscript{2}O,
1.3 mM KH\textsubscript{2}PO\textsubscript{4}, 24 mM NaHCO\textsubscript{3},
2.0 mM CaCl\textsubscript{2} $\cdot$ 2 H\textsubscript{2}O.
The needles used for the injections were pulled with a
Sutter Instrument Co. Model P-2000 from thin-walled glass
capillaries (1.00 mm; VWR) and cut open with a forceps.
A volume of 1 nl of the solution was injected with a pressure injector
(Eppendorf Femtojet 4i) in the rostral rhombencephalic ventricle.
The pressure and time were calibrated for each needle using a 0.01 mm calibration slide.

Following injection, the zebrafish was directly placed under the confocal
microscope and 1200 images were acquired at a frequency of 13.1 Hz.
A single optical section was obtained. Particle tracking was done using
TrackMate~\cite{Tinevez2017TrackMate:Tracking} in Fiji/ImageJ~\cite{Schindelin2012Fiji:Analysis}
and plotted using MATLAB, as described in Jeong~\emph{et al.}~\cite{Jeong2022MeasurementTelencephalon}.
The parameters for TrackMate were as follows.
The DoG detector was used with the settings: threshold: 2.0, with
median filtering; radius: 2.0, with subpixel localization.
Next, the Simple LAP tracker was used with settings: max frame gap: 2;
alternative linking cost factor: 1.05; linking max distance: 5.0; gap closing max distance: 2.0;
splitting max distance: 15.0; allow gap closing: true; allow track splitting:
false; allow track merging: false; merging max distance: 15.0; cutoff percentile: 0.9.
Only the particle tracks with at least 30 data points, while simultaneously
covering a distance of at least 8.8 \textmu m, were plotted in~\Crefpanels{fig:fig1}{b, c}.

>>>>

The larval and adult zebrafish were raised under standard husbandry
conditions at 28.5°C in a Techniplast Zebtech Multilinking system.
The fish tanks were kept at constant pH 7.0 and 685 \textmu S with a
14/10 hr light/dark cycle.
From fertilization to 3 days post-fertilization (dpf), larvae were maintained in egg
water (1.2 g marine salt and 0.1\% methylene blue in 20 L reverse osmosis (RO) water)
and subsequently transferred to artificial fish water (AFW)
(1.2 g marine salt in 20 L RO water). The zebrafish lines used in the experiments were
\emph{ccdc103(dnaaf19)\textsuperscript{tn222a}}
(\emph{schmalhans (smh)})~\cite{Jau-NianChen1997Left-rightZebrafish}
and \emph{Tg(gfap:Gal4FF)\textsuperscript{nw7Tg}}~\cite{DiazVerdugo2019Glia-neuronSeizures},
\emph{Tg(5xuas:Signal-Dendra2)\textsuperscript{nw21Tg}}.
Animals were in the pigmentless \emph{nacre\textsuperscript{b692}}
(\emph{mitfa\textsuperscript{-/-}})~\cite{JamesA.Lister1999NacreFate} background.

Wholemount in vivo live imaging and photoconversion
experiments were performed with zebrafish larvae at the 2 dpf stage obtained from inbreeding of heterozygous
\emph{smh\textsuperscript{+/-};Tg(gfap:Gal4FF);Tg(5xuas:Signal-Dendra2})
adult animals, as described by D'Gama~\emph{et al.}~\cite{DGama2024Cilia-mediatedBrain}.
Controls consisted of either wild-type (\emph{smh}\textsuperscript{+/+})
or \emph{smh}\textsuperscript{+/-} heterozygous zebrafish from the same breeding.
The mutants were identified based on their curved body~\cite{Jau-NianChen1997Left-rightZebrafish}.

For genotyping, genomic DNA (gDNA) was isolated from clipped fins of anesthetized adult
fish using 100 \textmu L of PCR lysis buffer (containing 1M tris pH 7--9, 0.5 M EDTA,
tritonX-100, and Proteinase K 0.1 mg/ml) overnight at 50°C.
To stop the lysis reaction, the samples were heated to 95°C for 10 minutes
and then centrifuged at 13000 rpm for 2 minutes.
The supernatant containing gDNA was utilized for KASP assays-based analysis.
The gDNAs were diluted (1:2) with water, and 3 \textmu L of the diluted gDNA was
used to perform the KASP assay following the guidelines of the manufacturer (LGC Biosearch Technologies™).
For each sample well, the \fixme{master} mix contained: 5 \textmu L of master mix; 0.14
\textmu L of assay mix; 1.86 \textmu L of RO water.

@@

\subsection{Wholemount zebrafish in vivo time-lapse imaging and photoconversion of Dendra2}
Two dpf larval zebrafish were anesthetized in 0.013\% MS222
(Ethyl 3-aminobenzoate methanesulfonate, Sigma) in AFW and
mounted laterally in 1.5\% low melting agarose in a Flurodish (VWR, FD35PDL-100),
to obtain a lateral view of the brain ventricles. After the
fish were positioned in the agarose, the dish stood for 5 min to allow the agarose to solidify.
After solidification, AFW containing 0.013\% MS222 was added,
and then the fish were transferred to the confocal microscope (LSM880 Examiner, Zeiss)
and imaged using a 20x water-immersion objective (Zeiss, NA 1.0, Plan-Apochromat)
at room temperature.

@@

\subsection{Photoconversion data analysis}
The image data analysis was performed in MATLAB and the codes are
openly available~\fixme{\cite{zenodo-link-to-all-code}}.  The
photoconverted channel was aligned to correct for drift in $x,y$
directions using previously developed
codes~\cite{Reiten2017Motile-Cilia-MediatedComputations,
  Ringers2023NovelEpithelia}.  Only stable recordings were further
analyzed.  To increase the signal-to-noise ratio, the images were
downsampled by a block average using a resampling factor of 6. % (using the MATLAB function blockproc).
The change in fluorescence intensity
$\Delta F = (F-F_0)/F_0$ was
calculated for each pixel and each time point.  Here, the
value $F_0$ is the average fluorescence intensity for the baseline
(the first 10 time-lapse images) acquired before photoconversion.

@@

To obtain $\Delta F$ values for each region of interest (ROI), ROIs were
first drawn on the aligned and block-averaged time series (\Cref{fig:fig2}).
Six ROIs were drawn manually: ROI 1 around the location of photoconversion,
and ROIs 2--6 in different regions of the ventricular system.
The pixel values of the fluorescence intensity within one ROI were
averaged for each time point, and the relative change $\Delta F$ calculated with this averaged value.
To normalize the data with respect to photoconversion efficiency, the
$\Delta F$ curves for all ROIs were divided by the $\Delta F$ values
obtained for ROI 1 at the end of the experiment,
so that $\Delta F$ values for the photoconverted site ROI 1 ranged from
zero to one. Finally, we averaged the $\Delta F$ data from controls and from mutants.
%We report these mean values together with a one-standard deviation error band.

\subsection{Intraventricular injection of microbeads and particle tracking}
%To recover the flow fields shown in~\Crefpanels{fig:fig1}{b, c},

We also carried out
particle tracking in a larval zebrafish injected with fluorescent beads, as described in previous work~\cite{Olstad2019CiliaryDevelopment}.
Anesthetized 2 dpf zebrafish were injected with 1 nl of a mixture containing 0.1\% w/v fluorescent beads
(SPHERO Fluorescent Yellow Particles 1\% w/v, F = 0.16 mm) diluted
in 7.5 mg/ml 70 kDa rhodamine B isothiocyanate-dextran (RITCdextran; Sigma-Aldrich, R9379)
dissolved in artificial CSF. Artificial CSF composition was as follows:
124 mM NaCl, 22 mM D-(+)-Glucose, 2.0 mM KCl, 1.6 mM MgSO\textsubscript{4} $\cdot$ 7 H\textsubscript{2}O,
1.3 mM KH\textsubscript{2}PO\textsubscript{4}, 24 mM NaHCO\textsubscript{3},
2.0 mM CaCl\textsubscript{2} $\cdot$ 2 H\textsubscript{2}O.
The needles used for the injections were pulled with a
Sutter Instrument Co. Model P-2000 from thin-walled glass
capillaries (1.00 mm; VWR) and cut open with a forceps.
A volume of 1 nl of the solution was injected with a pressure injector
(Eppendorf Femtojet 4i) in the rostral rhombencephalic ventricle.
The pressure and time were calibrated for each needle using a 0.01 mm calibration slide.

Following injection, the zebrafish was directly placed under the confocal
microscope and 1200 images were acquired at a frequency of 13.1 Hz.
A single optical section was obtained. Particle tracking was performed using
TrackMate~\cite{Tinevez2017TrackMate:Tracking} in Fiji/ImageJ~\cite{Schindelin2012Fiji:Analysis}
and plotted using MATLAB, as described by Jeong~\emph{et al.}~\cite{Jeong2022MeasurementTelencephalon}.
The parameters for TrackMate were as follows.
The DoG detector was used with the settings: threshold: 2.0, with
median filtering; radius: 2.0, with subpixel localization.
Next, the Simple LAP tracker was used with settings: max frame gap: 2;
alternative linking cost factor: 1.05; linking max distance: 5.0; gap closing max distance: 2.0;
splitting max distance: 15.0; allow gap closing: true; allow track splitting:
false; allow track merging: false; merging max distance: 15.0; cutoff percentile: 0.9.
Only the particle tracks with at least 30 data points, while simultaneously
covering a distance of at least 8.8 \textmu m, are shown (\Crefpanels{fig:fig1}{b, c}).
